Integration of Multiple Constraints in ACG

### 2014-02-18 lenls/post-proc.tex
POST: Final remarks from Maxime, flipped sigs in story diagrams

G-ACG extrinsic languages correspond to ACG object languages with
intersection. More formally, while ACG object languages are ACG abstract
languages closed on transformation by a lexicon, G-ACG extrinsic languages are
ACG abstract languages closed on transformation by a lexicon and
intersection\footnote{Can be shown by induction on the topological ordering of
  any given graphical ACG.}.

@@

@@

Pangraphical languages turn out to be at least as expressive as the extrinsic
languages\footnote{This might come as no surprise given the extended domain of
  constraints compared to extrinsic languages (constrained by the entire graph
  as opposed to just the node's predecessors).}. The proof is carried out by
transforming a G-ACG such that a particular node will have the same
pangraphical language in the latter G-ACG as the extrinsic language it had in
the former. The nodes of the newly constructed G-ACG correspond to paths in
the former one terminating in the node in question\footnote{This means that
  when the G-ACG is $\mathcal{D}$, the diamond grammar, and the node in
  question is $Bottom$, then we will have two different nodes for the two
  paths $[Top, Left, Bottom]$ and $[Top, Right, Bottom]$.}. Thus the labelling
of nodes with terms which certifies a given term's presence in the
pangraphical language of the new G-ACG also serves as a proof of its presence
in the extrinsic language of the old G-ACG and vice versa.

@@

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{0.4\textwidth}
    \centering
    \includegraphics[height=0.2\textheight]{../diagrams/single-cont.pdf}
    \caption{\label{fig:single-cont} An extraction constraint }
  \end{subfigure}
  \qquad
  \begin{subfigure}[b]{0.4\textwidth}
    \centering
    \includegraphics[height=0.2\textheight]{../diagrams/merged-conts.pdf}
    \caption{\label{fig:merged-conts} G-ACG merging the three extraction
      constraints.}
  \end{subfigure}
  \caption{\label{fig:conts} Combining the constraints on extraction.}
\end{figure}

@@

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{0.4\textwidth}
    \centering
    \includegraphics[height=0.2\textheight]{../diagrams/conts-elab.pdf}
    \caption{\label{fig:conts-elab} Elaborating the lexicon from syntax to
      strings. }
  \end{subfigure}
  \qquad
  \begin{subfigure}[b]{0.4\textwidth}
    \centering
    \includegraphics[height=0.2\textheight]{../diagrams/final-synth.pdf}
    \caption{\label{fig:final-synth} Adding our own syntactic constraints,
      $Neg$ and $Agr$.}
  \end{subfigure}
  \caption{\label{fig:final} Building the example grammar up to its final
    state.}
\end{figure}

@@

\begin{align*}
N_{aucun} &: N_{NEG{=}F} \limp NP_{NEG{=}T} \\
N_{le_1} &: N_{NEG{=}F} \limp NP_{NEG{=}F} \\
N_{le_2} &: N_{NEG{=}T} \limp NP_{NEG{=}T} \\
N_{ne^{tv}_1} &: (NP_{NEG{=}F} \limp NP_{NEG{=}F} \limp S) \limp (NP_{NEG{=}T} \limp NP_{NEG{=}F} \limp S) \\
N_{ne^{tv}_2} &: (NP_{NEG{=}F} \limp NP_{NEG{=}F} \limp S) \limp (NP_{NEG{=}F}
\limp NP_{NEG{=}T} \limp S) \\
N_{ne^{tv}_3} &: (NP_{NEG{=}F} \limp NP_{NEG{=}F} \limp S) \limp (NP_{NEG{=}T}
\limp NP_{NEG{=}T} \limp S) \\
N_{aime} &: NP_{NEG{=}F} \limp NP_{NEG{=}F} \limp S \\
N_{que_1} &: (NP_{NEG{=}F} \limp S) \limp N_{NEG{=}F} \limp N_{NEG{=}F} \\
N_{que_2} &: (NP_{NEG{=}F} \limp S) \limp N_{NEG{=}T} \limp N_{NEG{=}T} \\
\end{align*}

@@

\begin{align*}
A_{tatou} &: N_{NUM{=}SG} \\
A_{tatous} &: N_{NUM{=}PL} \\
A_{le} &: N_{NUM{=}SG} \limp NP_{NUM{=}SG} \\
A_{les} &: N_{NUM{=}PL} \limp NP_{NUM{=}PL} \\
A_{aime_1} &: NP_{NUM{=}SG} \limp NP_{NUM{=}SG} \limp S \\
A_{aime_2} &: NP_{NUM{=}SG} \limp NP_{NUM{=}PL} \limp S
\end{align*}
\begin{align*}
A_{qui_1} &: (NP_{NUM{=}SG} \limp S) \limp N_{NUM{=}SG} \limp N_{NUM{=}SG} \\
A_{qui_2} &: (NP_{NUM{=}PL} \limp S) \limp N_{NUM{=}PL} \limp N_{NUM{=}PL}
\end{align*}

>>>>

G-ACG extrinsic languages correspond to ACG object languages with
intersection. More formally, while ACG object languages are ACG abstract
languages closed on transformation by a lexicon, G-ACG extrinsic languages are
ACG abstract languages closed on transformation by a lexicon and
intersection.

@@

\begin{proof}
  We prove $\mathcal{E} = \mathcal{A}^{\mathcal{L}{\cap}}$, or equivalently
  $\mathcal{E} = \mathcal{I}^{\mathcal{L}{\cap}}$, by proving $\mathcal{E}
  \subseteq \mathcal{I}^{\mathcal{L}{\cap}}$ and
  $\mathcal{I}^{\mathcal{L}{\cap}} \subseteq \mathcal{E}$.

  We begin with $\mathcal{E} \subseteq \mathcal{I}^{\mathcal{L}{\cap}}$. Let
  us consider any graphical ACG $\mathcal{G}$ and some topological ordering
  $v_1, \ldots, v_n$ of its nodes $V(G)$ (an ordering such that more abstract
  nodes always precede less abstract ones). We proceed by induction on the
  topological ordering and prove that $\mathcal{E}_{\mathcal{G}}(v_i) \in
  \mathcal{I}^{\mathcal{L}{\cap}}$ for all $v_i$ in $V(G)$:

  \begin{itemize}
  \item For $v_1$, the case is trivial. $v_1$ has no predecessors and so
    $\mathcal{E}_{\mathcal{G}}(v_1) = \mathcal{I}_{\mathcal{G}}(v_1) \in
    \mathcal{I}$.
  \item For any other node $v_n$, we inspect the definition of
    $\mathcal{E}_{\mathcal{G}}(v_n)$.

    $$\mathcal{E}_{\mathcal{G}}(v_n) = \mathcal{I}_{\mathcal{G}}(v_n) \cap
    \bigcap_{(u,v_n) \in E(G)}
    \mathcal{L}_{(u,v_n)}(\mathcal{E}_{\mathcal{G}}(u))$$

    By induction hypothesis, $\mathcal{E}_{\mathcal{G}}(u) \in
    \mathcal{I}^{\mathcal{L}{\cap}}$ for all the predecessors $u$ of
    $v_n$. Since, $\mathcal{I}^{\mathcal{L}{\cap}}$ is closed on projection by
    a lexicon, we have that
    $\mathcal{L}_{(u,v_n)}(\mathcal{E}_{\mathcal{G}}(u)) \in
    \mathcal{I}^{\mathcal{L}{\cap}}$ as well. Finally, by the closure of
    $\mathcal{I}^{\mathcal{L}{\cap}}$ on intersection, we have that their
    intersection $\bigcap_{(u,v_n) \in E(G)}
    \mathcal{L}_{(u,v_n)}(\mathcal{E}_{\mathcal{G}}(u))$ also belongs to
    $\mathcal{I}^{\mathcal{L}{\cap}}$. Since $\mathcal{I}_{\mathcal{G}}(v_n)$
    belongs to $\mathcal{I}^{\mathcal{L}{\cap}}$ too, we know that their
    intersection, $\mathcal{E}_{\mathcal{G}}(v_n)$ belongs to
    $\mathcal{I}^{\mathcal{L}{\cap}}$.
  \end{itemize}

  Now we prove the converse, that is that $\mathcal{I}^{\mathcal{L}{\cap}}
  \subseteq \mathcal{E}$. We do so by proving that $\mathcal{I} \subseteq
  \mathcal{E}$ and that $\mathcal{E}$ is closed on projection through a
  lexicon and intersection.

  \begin{description}
    \item[Inclusion of intrinsic languages] Let us consider the intrinsic
      language $\mathcal{I}_{\mathcal{G}}(u)$ of some node $u$ in some G-ACG
      $\mathcal{G}$. We can produce a G-ACG $\mathcal{G}_u$, whose graph
      contains only the node $u$ labeled with the same signature and
      distinguished type as in $\mathcal{G}$. Since $u$ has no predecessors in
      $\mathcal{G}_u$, it will be the case that
      $\mathcal{E}_{\mathcal{G}_u}(u) = \mathcal{I}_{\mathcal{G}_u}(u) =
      \mathcal{I}_{\mathcal{G}}(u)$ and therefore the intrinsic language
      $\mathcal{I}_{\mathcal{G}}(u)$ is also an extrinsic language
      $\mathcal{E}_{\mathcal{G}_u}(u)$.

    \vspace{5mm}

    \item[Closure on projection through a lexicon] Let $L_1$ and $L_2$ be two
      extrinsic languages defined by the nodes $u_1$ and $u_2$ in some
      graphical ACGs $\mathcal{G}_1$ and $\mathcal{G}_2$, respectively. We can
      construct a new graphical ACG $\mathcal{G}'$ by taking the union of the
      two ACGs (union of graphs (union of vertices and edges) and unions of
      the labelings).\footnote{Assuming, without loss of generality, that the
        sets of vertices of the two graphs are disjoint.} We will then add a
      new node $v$ such that $\mathcal{E}_{\mathcal{G}'}(v) = L_1 \cap L_2$.

      The signature labeling $v$ will be the union of the two signatures
      labeling $u_1$ and $u_2$, ensuring that $\forall u \in \{u_1,
      u_2\}$, $\mathcal{I}_{\mathcal{G}'}(v) \supseteq
      \mathcal{I}_{\mathcal{G}'}(u) \supseteq
      \mathcal{E}_{\mathcal{G}'}(u)$. We can use either of the two
      distinguished types that label $u_1$ or $u_2$ to label $v$ (if
      they are different, then the intersection $L_1 \cap L_2$ is
      trivially empty). Finally, we will add two edges leading from
      $u_1$ and $u_2$ to $v$, both labeled with the identity lexicon. We
      will now show that $\mathcal{E}_{\mathcal{G}'}(v) = L_1 \cap L_2$.

      \begin{align*}
        \mathcal{E}_{\mathcal{G}'}(v) &= \mathcal{I}_{\mathcal{G}'}(v)
        \cap \mathcal{L}_{(u_1,v)}(\mathcal{E}_{\mathcal{G}'}(u_1)) \cap
        \mathcal{L}_{(u_2,v)}(\mathcal{E}_{\mathcal{G}'}(u_2)) \\ &=
        \mathcal{I}_{\mathcal{G}'}(v) \cap
        \mathcal{E}_{\mathcal{G}'}(u_1) \cap
        \mathcal{E}_{\mathcal{G}'}(u_2) \\ &=
        \mathcal{E}_{\mathcal{G}'}(u_1) \cap
        \mathcal{E}_{\mathcal{G}'}(u_2) \\ &=
        \mathcal{E}_{\mathcal{G}_1}(u_1) \cap
        \mathcal{E}_{\mathcal{G}_2}(u_2) \\ &= L_1 \cap L_2
      \end{align*}

      The simplifications use the facts that: the two new lexicons are
      identities, that $\mathcal{I}_{\mathcal{G}'}(v) \supseteq
      \mathcal{E}_{\mathcal{G}'}(u_1) \cap \mathcal{E}_{\mathcal{G}'}(u_2)$
      (due to $\Sigma'_v = \Sigma'_{u_1} \cup \Sigma'_{u_2}$), that
      $\mathcal{E}_{\mathcal{G}'}(u_1) = \mathcal{E}_{\mathcal{G}_1}(u_1)$ and
      $\mathcal{E}_{\mathcal{G}'}(u_2) = \mathcal{E}_{\mathcal{G}_2}(u_2)$
      (due to $u_1$ and $u_2$ having the same predecessors in $\mathcal{G}$
      and $\mathcal{G}'$) and finally that $\mathcal{E}_{\mathcal{G}}(u_1) =
      L_1$ and $\mathcal{E}_{\mathcal{G}}(u_2) = L_2$ (our starting
      assumptions).

  \vspace{5mm}

  \item[Closure on intersection] Let $L \subseteq \Lambda(\Sigma_O)$ be a
    language which is the result of mapping an extrinsic language $L'
    \subseteq \Lambda(\Sigma_A)$ through the lexicon $\mathcal{L} : \Sigma_A
    \to \Sigma_O$. We will show that $L$ is also an extrinsic language. Let
    $u$ be the node that defines the extrinsic language $L$ in some graphical
    ACG $\mathcal{G}$. We construct a new graphical ACG $\mathcal{G}'$ by
    adding a new vertex $v$ to $\mathcal{G}$ labeled with the signature
    $\Sigma_O$ and the distinguished type $\mathcal{L}(S)$ where $S$ is the
    distinguished type labeling $u$. To this graph, we conjoin the edge
    $(u,v)$ and label it with $\mathcal{L}$.

    Now we verify that $\mathcal{E}_{\mathcal{G}'}(v) = L$.

    \begin{align*}
      \mathcal{E}_{\mathcal{G}'}(v) &= \mathcal{I}_{\mathcal{G}'}(v) \cap
      \mathcal{L}(\mathcal{E}_{\mathcal{G}'}(u)) \\
      &= \mathcal{L}(\mathcal{E}_{\mathcal{G}'}(u)) \\
      &= \mathcal{L}(\mathcal{E}_{\mathcal{G}}(u)) \\
      &= \mathcal{L}(L') \\
      &= L
    \end{align*}

    The assumptions used here are in turn: that $\mathcal{I}_{\mathcal{G}'}(v)
    \supseteq \mathcal{L}(\mathcal{E}_{\mathcal{G}'}(u))$ (due to the
    homomorphism property of lexicon $\mathcal{L} : \Sigma_A \to \Sigma_O$),
    that $\mathcal{E}_{\mathcal{G}'}(u) = \mathcal{E}_{\mathcal{G}}(u)$ (due
    to $u$ having the same predecessors in $\mathcal{G}$ and $\mathcal{G}'$),
    that $\mathcal{E}_{\mathcal{G}}(u) = L'$ and that $\mathcal{L}(L') = L$
    (our starting assumptions).
  \end{description}
\hfill $\qed$
\end{proof}

@@

Pangraphical languages turn out to be at least as expressive as the extrinsic
languages\footnote{This might come as no surprise given the extended domain of
  constraints compared to extrinsic languages (constrained by the entire graph
  as opposed to just the node's predecessors).}. The proof is carried out by
transforming any G-ACG $\mathcal{G}$ where $L = \mathcal{E}_{\mathcal{G}}(u)$
to a G-ACG $\mathcal{G'}$ where for some $v$, $\mathcal{P}_{\mathcal{G'}}(v) =
L$. We do so by taking all the (finitely many) paths in $G$ that end in $u$ to
be the vertices of $G'$. Each of the vertices in $G'$ (paths in $G$) will be
labeled by the same signature and distinguished type as the first node of the
corresponding path. Furthermore, there will be an edge $(p_1, p_2)$ in $G'$
labeled with the lexicon $\mathcal{L}$ whenever the path $p_2$ is the maximal
strict suffix of $p_1$ such that $p_1 = e; p_2$ where the edge $e$ is labeled
with $\mathcal{L}$ in $G$.\footnote{This means that when the G-ACG is
  $\mathcal{D}$, the diamond grammar, and the node in question is $Bottom$,
  then we will have the following nodes in $\mathcal{D}'$: $[Top, Left,
    Bottom]$, $[Top, Right, Bottom]$, $[Left, Bottom]$, $[Right, Bottom]$,
  $[Bottom]$. Note that we have two nodes for (i.e. beginning with) $Top$,
  which will let us admit two different antecedents in
  $\mathcal{E}_{\mathcal{D}}(Top)$.} In this G-ACG $\mathcal{G'}$, valid
models $T$ which assign $T([u]) = t$ are isomorphic to proofs of membership of
$t$ in $\mathcal{E}_{\mathcal{G}}(u)$ and therefore
$\mathcal{P}_{\mathcal{G'}}([u]) = \mathcal{E}_{\mathcal{G}}(u)$.

@@

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{0.4\textwidth}
    \centering
    \includegraphics[height=0.2\textheight]{../diagrams/story-single-cont.pdf}
    \caption{\label{fig:single-cont} An extraction constraint }
  \end{subfigure}
  \qquad
  \begin{subfigure}[b]{0.4\textwidth}
    \centering
    \includegraphics[height=0.2\textheight]{../diagrams/story-merged-conts.pdf}
    \caption{\label{fig:merged-conts} G-ACG merging the three extraction
      constraints.}
  \end{subfigure}
  \caption{\label{fig:conts} Combining the constraints on extraction.}
\end{figure}

@@

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{0.4\textwidth}
    \centering
    \includegraphics[height=0.2\textheight]{../diagrams/story-conts-elab.pdf}
    \caption{\label{fig:conts-elab} Elaborating the lexicon from syntax to
      strings. }
  \end{subfigure}
  \qquad
  \begin{subfigure}[b]{0.4\textwidth}
    \centering
    \includegraphics[height=0.2\textheight]{../diagrams/story-final-synth.pdf}
    \caption{\label{fig:final-synth} Adding our own syntactic constraints,
      $Neg$ and $Agr$.}
  \end{subfigure}
  \caption{\label{fig:final} Building the example grammar up to its final
    state.}
\end{figure}

@@

\begin{align*}
N_{aucun} &: N_{NEG{=}F} \limp NP_{NEG{=}T} \\
\end{align*} % for page break purposes
\begin{align*}
N_{le_1} &: N_{NEG{=}F} \limp NP_{NEG{=}F} \\
N_{le_2} &: N_{NEG{=}T} \limp NP_{NEG{=}T} \\
N_{ne^{tv}_1} &: (NP_{NEG{=}F} \limp NP_{NEG{=}F} \limp S) \limp (NP_{NEG{=}T} \limp NP_{NEG{=}F} \limp S) \\
N_{ne^{tv}_2} &: (NP_{NEG{=}F} \limp NP_{NEG{=}F} \limp S) \limp (NP_{NEG{=}F}
\limp NP_{NEG{=}T} \limp S) \\
N_{ne^{tv}_3} &: (NP_{NEG{=}F} \limp NP_{NEG{=}F} \limp S) \limp (NP_{NEG{=}T}
\limp NP_{NEG{=}T} \limp S) \\
N_{aime} &: NP_{NEG{=}F} \limp NP_{NEG{=}F} \limp S \\
N_{que_1} &: (NP_{NEG{=}F} \limp S) \limp N_{NEG{=}F} \limp N_{NEG{=}F} \\
N_{que_2} &: (NP_{NEG{=}F} \limp S) \limp N_{NEG{=}T} \limp N_{NEG{=}T} \\
\end{align*}

@@

\begin{align*}
A_{tatou} &: N_{NUM{=}SG} \\
A_{tatous} &: N_{NUM{=}PL} \\
A_{le} &: N_{NUM{=}SG} \limp NP_{NUM{=}SG} \\
A_{les} &: N_{NUM{=}PL} \limp NP_{NUM{=}PL} \\
A_{aime_1} &: NP_{NUM{=}SG} \limp NP_{NUM{=}SG} \limp S \\
A_{aime_2} &: NP_{NUM{=}SG} \limp NP_{NUM{=}PL} \limp S \\
\end{align*} % for page break purposes
\begin{align*}
A_{qui_1} &: (NP_{NUM{=}SG} \limp S) \limp N_{NUM{=}SG} \limp N_{NUM{=}SG} \\
A_{qui_2} &: (NP_{NUM{=}PL} \limp S) \limp N_{NUM{=}PL} \limp N_{NUM{=}PL} \\
\end{align*}
